---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nmmr

## Installation

nmmr relies on the rstan interface to [Stan](https://mc-stan.org). To install nmmr, first [follow instructions for setting up rstan](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started).

After successfully installing rstan, nmmr can be installed with devtools

```{r, install, eval=FALSE}
# install.packages(devtools)
library(devtools)
devtools::install_gitlab("psadil/nmmr")
```

# Example Analysis

## Data

A sample dataset is provided with this package. The dataset contains the beta values for a single participant, and shows the format expected by the functions of this package. The dataset can be loaded with the following command.

```{r, data}
# library(nmmr)
data("sub02")
knitr::kable(head(sub02))
```

For extra info on the dataset, see the help page for betas (?betas).

## Run Stan

The technique works by comparing separate models, each of which allows just a single kind of modulation to the neural tuning functions. The two kinds of neuromodulation currently implemented are _Additive_ and _Multiplicative_. Source for the models can be found on this package's repository. The three models are largely the same, differing only slightly in the NTFs for the high contrast.

### Define Stan options

In this simple example, most voxel-wise parameters are not saved (e.g., the weights for each channel in each voxel, the value of the modulation parameter for each voxel). Excluding these parameters drastically reduces the size of the output and speeds up post-processing. The parameter _mu_, which is the distribution of the beta values for each trial, might also be worth dropping. _mu_ is kept here because it is required for model comparison.

A few additional parameters are used to control Stan's sampling behavior. See the help page for rstan::stan. Running one chain may require a few hours.


