---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nmmr

Experimental package for NeuroModulation Modeling in R. This package under active development and not yet released.

## Installation

`nmmr` relies on the [RStan](https://github.com/stan-dev/rstan) interface to [Stan](https://mc-stan.org). To install `nmmr`, first [follow instructions for setting up RStan](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started).

After installing `RStan`, `nmmr` can be installed with the `remotes` package.

```{r, install, eval=FALSE}
# install.packages(remotes)
library(remotes)
devtools::install_github("psadil/nmmr")
```

# Example Analysis

## Data

The package is bundled with a sample dataset. The dataset contains the beta values for a single participant and shows the format expected by the functions of this package. The dataset can be loaded with the following command.

```{r, data}
# library(nmmr)
data("sub02")
knitr::kable(head(sub02))
```

For extra info on the dataset, see the help page for sub02 (?sub02).
